Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-06 paper/paper.tex
a small pass

\begin{mdframed}
  \textbf{Time-invariance assumption:} It is assumed that the MDP has reached steady-state conditions so that
  the dynamics of the process are equivalent over sufficiently long time windows of equal length.
\end{mdframed}

@@

\subsection{The Q-value: A goodness measure for action choices}
We can now formulate the purpose of the DMN
as finding a policy $\pi$ for the human agent that maximizes the
expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
\end{equation}
In other words,
the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$,
and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior,
potentially driven by the DMN in human agents.
% The goal of the DMN is then
% to find then to produce a policy $\pi$ which maximizes the expected cumulative reward $Q^\pi(s_0,\pi(a_0))$
% given an initial state $s_0$.

@@

\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992) generates a
\textit{greedy policy}
\textbf{XXX please add 1-2 sentences why this is interesting in the current
context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
  \end{split}
  \label{eq:bellman}
\end{equation}
This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function \textbf{XXX please more explicitly introduce
the notion of value function}
into the immediate reward
and the discounted rewards of the upcoming states.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in \eqref{eq:bellman} the sampling is done only over things which
depend on the environment, and so can be learned off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.

>>>>

\begin{mdframed}
  \textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary --i.e has reached steady-state conditions--
  so that the dynamics of the process are equivalent over sufficiently long time windows of equal length.
\end{mdframed}

@@

\subsection{The Q-values, state-values, and the advantage function}
We can now formulate the purpose of the DMN
as finding a policy $\pi$ for the human agent that maximizes the
expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a].
\end{equation}
In other words,
the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$,
and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior,
potentially driven by the DMN in human agents.

One can also define the state-value function

\begin{eqnarray}
  \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
\end{eqnarray}
which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
as a baseline value for any action. Finally, the advantage function is defined by
\begin{equation}
A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
\end{equation}

% The goal of the DMN is then
% to find then to produce a policy $\pi$ which maximizes the expected cumulative reward $Q^\pi(s_0,\pi(a_0))$
% given an initial state $s_0$.

@@

\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992) generates a
\textit{greedy policy}
\textbf{XXX please add 1-2 sentences why this is interesting in the current
context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) = \argmax_{a \in \mathcal A}{Q}^*(s, a),
  \label{eq:qlearning}
\end{equation}
which maps each state $s$ to the optimal action which should be taken at that state.
Here $Q^*(s, a) := \max_{\pi}Q^\pi(s,a)$, the optimal Q-value over all policies.
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^*(s_0, a_0) &= \mathbb E_{s_1 \sim p(s_1|s_0,a)} [r(s,a) + \gamma \max_{a_1 \in \mathcal A}Q^*(s_1, a_1)]
    \\
    &= \mathbb E_{s_1 \sim p(s_1|s_0,a_0)} [r(s_0,a_0) + \gamma Q^*(s_1, \pi(s_1))].
  \end{split}
  \label{eq:bellman}
\end{equation}
This consistency equation for the optimal (over all policies) Q-value function provides
a recursive decomposition of optimal behavior by dividing the Q-value function into the immediate
reward and the discounted rewards of the upcoming states. The optimal Q-value operator $Q^*$
is a fixed point for this equation. A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in \eqref{eq:bellman} the sampling is done only over things which
depend on the environment, and so can be learned off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.
